### 2018-03-16 chapter2/ms2.tex
Fixing Ch 2. Needs more proofreading

%UNLIKE IN A REGULAR TEX FILE, DON'T PUT ANY PREAMBLE MATERIAL HERE

@@

\clearpage
Extensive theoretical and observational work has been performed to understand the potential EM emission from a merging compact object binary. The range of potential counterparts covers the entire EM spectrum. The most popular counterpart is a short-hard gamma-ray burst \citep[SGRB; e.g.,][]{Paczynski1986,Narayan+92,Berger2014} powered by rapid accretion onto the compact remnant which leads to a strongly beamed relativistic jet. The evidence for a connection between SGRBs and compact object mergers includes the association with older stellar populations than for long GRBs \citep{Berger+05,Bloom+06,Berger2011,Fong+11,Fong+13}, large physical offsets between the burst locations and their host galaxies indicative of progenitor kicks \citep{fong+10, FongBerger13}, weak association with the underlying stellar light of their hosts \citep{fong+10, FongBerger13}, and a lack of association with supernovae \citep{fox+05,Soderberg+06,Berger2010,FongBerger13,Berger2014}. The relativistic jet can also be observed in the x-ray, optical, and radio as non-thermal afterglow emission arises from interaction with the ambient medium. This emission is also affected by relativistic beaming but may still be seen off-axis as the jet decelerates and expands. In particular, for large off-axis angles $(\theta\gtrsim 2 \theta_j)$ the emission will be most pronounced in the radio on timescales of weeks to years \citep{NakarPiran11,MetzgerBerger12}.

@@

To investigate the detectability of kilonovae, and to asses potential contamination in deep high-cadence, wide-field searches for such sources, we carried out a set of observations using the Dark Energy Camera (DECam) mounted on the CTIO/Blanco 4m telescope (PI: Berger). This GW follow-up ``dry-run" covered an area typical for a GW error region ($\sim$ 70 deg$^2$), with two visits per night for five nights to a depth of $\approx$ 24 AB mag in both {\em i}- and {\em z}-band. This rich data set will serve as the framework to study the nature of foreground/background contamination and to test the effectiveness of various observing strategies and contaminant rejection strategies when performing follow-up observations of a GW event.

@@

\clearpage
\section{Kilonovae and Contaminants}
\label{sec:ch2_contaminants}

@@

A critical property for understanding kilonova emission is the opacity, as the ability of photons to diffuse through the ejecta determines the timescale, peak luminosity, and the spectral energy distribution of the transient. \citet{Metzger+10}, using the models of \citep[LP98 hereafter]{LP98}, found that a typical Fe-peak element opacity $(\kappa \sim 0.1 \; \text{cm}^2 \text{ g}^{-1} \text{ for } \lambda \sim 0.3 - 1\; \mu\text{m})$ leads to a peak time of $t_p \sim 1$ day, with a bolometric peak luminosity of $L_p \sim 10^{41} \text{ erg s}^{-1}$, and a photospheric temperature of $T_p \approx 10^4$ K (i.e., a peak in the UV). However, subsequent work by \citet{Kasen+13} and \citet{BarnesKasen13} found that {\em r}-process material have much larger opacities with $\kappa \sim 10-100 \; \text{cm}^2 \text{ g}^{-1} \text{ for } \lambda \sim 0.3 - 3\; \mu\text{m}$. This leads to a transient that is longer lived $(t_p \sim$ days), fainter $(L_p \sim 10^{40}$ erg s$^{-1})$, and much redder $(T_p \approx 3000$ K, e.g., a peak in the NIR). These results have been confirmed in simulations by other groups \citep[see e.g.,][]{TanakaHotokezaka13,Grossman+14,Tanaka+14}.

In this work, we construct kilonova light curves using the time-resolved spectra of \citet{BarnesKasen13}. They consider a grid of models with $\Mej = 10^{-3}\;, 10^{-2} \;, \text{ and } 10^{-1} \;M_{\odot}$ and with $\beta_{\text{ej}} = $ 0.1, 0.2, and 0.3. Unless otherwise noted we only consider the average ejecta properties of $\Mej =10^{-2}\;M_{\odot}$ and $\beta_{\text{ej}} = 0.2$ since we do not know the distribution of ejecta properties a priori. We compute light curves using the DECam filter transmission curves\footnote{\url{http://www.ctio.noao.edu/noao/sites/default/files/DECam/DECam_filters_transmission.txt}}\footnote{We note that the DECam transmission filters are fairly standard so these results are relevant for other instruments as well.} with the luminosity $L_{\phi}$, through a filter $\phi$, at a time t, being given by
\begin{align}
\label{eq:LCint}
L_{\phi}(t) = \frac{\int d\lambda L_{\lambda}(t,\lambda) \phi(\lambda)}{\int d\lambda \phi(\lambda)}
\end{align}
\noindent where $L_{\lambda}(t,\lambda)$ is the model specific luminosity and $\phi(\lambda)$ is the filter transmission as a function of wavelength. A representative plot of the spectra at 1 day and 1 week, along with the resulting {\em r}-, {\em i}-, and {\em z}-band light curves are shown in \Cref{fig:ch2_kilosample}.

\begin{figure}[t!]
\centering
\includegraphics[width=\textwidth]{./figs/chapter2/ch2_f1.pdf}
\caption{ \singlespace Top: Kilonova spectra from \citet{BarnesKasen13} at $\sim$ 1 d and $\sim$ 1 week with ejecta properties of $\Mej =10^{-2}\;M_{\odot}$ and $\beta_{\text{ej}} = 0.2$. Bottom: Light curve for the same model parameters computed at a fiducial distance of 200 Mpc in the DECam {\em r}, {\em i}- and {\em z}-band filters. \vspace{10pt}}
\label{fig:ch2_kilosample}
\end{figure}

@@

\clearpage
\begin{figure}[h!]
\centering
\includegraphics[width=0.65\textwidth]{./figs/chapter2/ch2_f2.pdf}
\caption{ \singlespace Top: {\em i}-band light curves of the contaminants considered in this work at a fiducial distance of 200 Mpc. Bottom: {\em z}-band light curves. See \cref{sec:ch2_WDmerge,sec:ch2_AIC,sec:ch2_ELDD,sec:ch2_type.Ia,sec:ch2_PSFast,sec:ch2_typeIa} for details.}
\label{fig:ch2_LCcont}
\end{figure}

@@

The slow evolution of the CSDD model light curve relative to kilonovae means that it can be easily rejected as a contaminant on the basis of timescales alone. However, we still consider the ELDD scenario as the more rapid rise time and faster decline means that it could serve as a potential contaminant. We make use of bolometric light curves from the simulations of \citet{Sim+12} to compute DECam {\em i}- and {\em z}-band light curves as in \cref{sec:ch2_WDmerge}. The ELDD spectrum is approximated as a blackbody with a temperature of $T_p \approx 4800$ K using the model spectra of \citet[see their Figure 7]{Sim+12}. The bolometric corrections are found to be $m_{\text{bol}} - m_{\text{i}} \approx -1.7$  mag and $m_{\text{bol}} - m_{\text{z}} \approx -1.9$ mag. We assume that these corrections are constant in time and apply them to the bolometric light curves to generate $i$- and $z$-band light curves (\Cref{fig:ch2_LCcont}).

@@

\clearpage
\subsection{Stellar Flares}
\label{sec:ch2_flarestar}
Stellar flares, primarily from M dwarfs, are another potential contaminant in fast cadence wide-field optical searches \citep{Becker+04,KulkarniRau06,Berger+13}.  These flares have characteristic timescales of minutes to hours \citep[see e.g.,][]{Berger+13} and temperature of $T_p \sim 10^4$ K, leading to an appearance as fast and blue transients.  As a result, M dwarf flares will only appear as a significant contaminant in kilonova searches that rely on single-epoch detection, primarily in $g$- and $r$-band in which the kilonova timescale is shorter than in the redder bands.  We return to this point in \cref{sec:ch2_disc}.  The sky-projected M dwarf flare rate at the depths relevant for kilonova searches is large, $\sim 10$ deg$^{-2}$ day$^{-1}$ \citep{KulkarniRau06}.  On the other hand, as was demonstrated by \citet{Berger+13}, deep observations in the $i$- and $z$-band can easily reveal the underlying M dwarfs and hence reduce the potential for contamination from flares.

\subsection{AGN Variability}
\label{sec:ch2_agns}
We also consider short-duration optical variability from active galactic nuclei (AGNs). AGN typically show strong variability on month or year-long timescales, making AGN outbursts too long in duration to be confused with a kilonovae. However, recent work has detected rare cases of day-long  or shorter variability in the NIR. \citet{Genzel+03} observed an H-band flare in Sgr A$^{*}$ during which the source brightened by a factor of 5. This flare lasted $\sim$ 30 min, with a characteristic rise time of $\sim2-5$ min. Such short timescale flares can be ruled out as kilonova contaminants by requiring a detection on more than a single night. Flares with longer durations are also unlikely to be significant contaminants. \citet{Totani+05} performed {\em V}- and {\em i}-band observations of six AGN, selected on the basis of their variability, over a period of four days. They found typical variability of only $\sim1-5\%$ above their host galaxy luminosity. More recent work by \citet{Ruan+12} studied the variability of a sample of \fermi selected blazars. They found that these sources showed a typical rms variability of $\sim0.6$ mag on timescales of $\sim 10$ d. Therefore, AGN are unlikely to exhibit sufficiently large variability to be confused with a kilonova.

\subsection{Volumetric Rates}
\label{sec:ch2_rates}
An additional important factor for understanding the impact of potential contaminants on fast cadence, wide-field searches is their volumetric rates. We note that the rates for these types of objects are generally poorly understood due to the lack of strong observational constraints; the notable exception are Type Ia SNe. The volumetric rates of potential contaminants are compiled in \cref{tab:ch2_rates}. No values are provided in the literature for the ELDD volumetric rate so we assign a fiducial value of 1\% of the Type Ia SN rate, consistent with the other rapid transients in this study. The volumetric rate for Type Ia SNe is known to be a function of redshift, and we use an approximate average volumetric rate in redshift bins of $\Delta z \sim 0.25$ out to $z_{\text{max}} \sim 0.75$. \citep[][and references therein]{Li+11,Graur+14}. These rates are listed in \cref{tab:ch2_rates_Iae}.

\clearpage
\singlespace
\begin{deluxetable}{cccccc}
\tabletypesize{\footnotesize}
\tablecolumns{6}
\tabcolsep0.05in\footnotesize
\tablewidth{0pt}
\tablecaption{Expected Rates For Fast Transients}
\tablehead{
\colhead{Object}&  \colhead{$\mathcal{R}_{\rm vol}$}& \colhead{$z_{\rm max}$}  &  \colhead{$\mathcal{R}_{\rm area}$} &  \colhead{$N$}     &     \colhead{Reference}      \\
\colhead{} & \colhead{(Mpc$^{-3}$ yr$^{-1}$)} & \colhead{} & \colhead{(deg$^{-2}$ yr$^{-1}$)} &  \colhead{} & \colhead{} }
\startdata
 Type .Ia    &                  10$^{-6}$                   &  0.5        &  $0.7$            &         1.4  & \citet{Bildsten+07} \\
 WD-NS mergers  &                  10$^{-5}$                   &   0.06        & $2\times10^{-2}$              &         $3\times10^{-2}$  &     \citet{Thompson+09}     \\
 WD-BH mergers  &      10$^{-5}$                   &    0.06        &   $2\times10^{-2}$              &      $3\times10^{-2}$  &    \citet{fryer+99}          \\
AIC       &                  10$^{-6}$                   & 0.06        & $2\times10^{-3}$            &        $3\times10^{-3}$ &  \citet{Darbha+10}   \\
ELDD      &                  3$\times 10^{-7}$                   & 0.14        &    $6\times10^{-3}$           &      $1\times10^{-2}$ &  $\cdots$  \\
Pan-STARRS fast &                  5$\times 10^{-6}$                   &     0.5        &   $3.5$            &     7 &   \citet{Drout+14}
\enddata
\tablecomments{Expected rates for the various contaminants considered in \cref{sec:ch2_contaminants}. $\mathcal{R}_{\rm area}$ is computed assuming an isotropic distribution of sources in a volume defined by the comoving volume at $z_{\rm max}$.  The column $N$ refers to the number of events expected during a search covering 100 deg$^2$ for 7 days. See \cref{sec:ch2_rates} or details.}
\label{tab:ch2_rates}
\end{deluxetable}

\clearpage
\begin{deluxetable}{cccc}
\tabletypesize{\footnotesize}
\tablecolumns{4}
\tabcolsep0.05in\footnotesize
\tablewidth{0pt}
\tablecaption{Expected Rates For Type Ia SNe}
\tablehead{
\colhead{Redshift}     &  \colhead{$\mathcal{R}_{\rm vol}$}&  \colhead{$\mathcal{R}_{\rm area}$}  &     \colhead{$N$} \\
\colhead{} & \colhead{(Mpc$^{-3}$ yr$^{-1}$)} &  \colhead{(deg$^{-2}$ yr$^{-1}$)} &  \colhead{} }
\startdata
0 -- 0.25 & $3\times10^{-5}$ & 3.5 & 21  \\
0.25 -- 0.50 & $4\times10^{-5}$ & 25 & 153  \\
0.5 -- 0.75 & $5\times10^{-5}$ & 65 & 394
\enddata
\tablecomments{Expected rates for Type Ia SNe in redshift bins of $\Delta z = 0.25$ out to $z_{\rm max} = 0.75$. See \cref{sec:ch2_rates}, \citet{Graur+14}, and references therein.}
\label{tab:ch2_rates_Iae}
\end{deluxetable}

\clearpage
\doublespace 
To estimate the expected number of events observed during a GW optical follow-up search we compute the number of sources per square degree per year. This is done by determining the maximum redshift to which the object can be observed to a limiting magnitude of $z \approx$ 23 mag (see \cref{sec:ch2_kilo,sec:ch2_MCsims_det}). This rate is then multiplied by the search area $(A \sim 100$ deg$^2$) and duration to get the number of expected sources for our observations. We compute the duration by assuming that a kilonova will be detected within approximately one day of the GW trigger. Therefore, we do not consider events that begin more than +2 d into the search. However, we do consider sources that begin up to twice their characteristic timescale ($\sim5$ d for rapid transients and $\sim 20$ d for Type Ia SNe) prior to the GW trigger, as they may remain visible when the search begins. The total duration is then 7 d for rapid transients and 22 d for Type Ia SNe. The number of expected sources is given in \cref{tab:ch2_rates,tab:ch2_rates_Iae}. Type Ia SNe dominate the number of expected events with $N \sim 568$ when summing across all redshift bins. Of the population of fast transients, only type .Ia and the Pan-STARRS fast transients are expected to have $N \gtrsim 1$. Despite the low expected occurrence rate for other fast transients, we still consider these sources in this analysis due to the large uncertainty in their rates.

@@

\begin{figure*}[t!]
\centering
\includegraphics[trim= 0 0 0 250 ,clip,width=\textwidth]{./figs/chapter2/ch2_f3.pdf}
\caption{ \singlespace Time-color (left) and time-magnitude (right) slices of the time-color-magnitude phase-space at a selection of redshifts. Arrow markers denote upper limits on the rise time. Note that in all panels the population of kilonovae are kept at the fiducial distance of $\sim$ 200 Mpc while redshift corrections are applied to the contaminant population. Kilonovae are well separated from the contaminant population in both color and timescales. The delineation in magnitude is less apparent.}
\label{fig:ch2_phase}
\end{figure*}

Investigating the first panel in the rise time-color space, we can see that the nine kilonovae models are well separated from the population of contaminating sources by virtue of their redder color. We also note that their rise times are generally shorter, with only the AIC and $1.0+0.05\;M_{\odot}$ type .Ia light curves exhibiting comparable timescales. The region occupied by kilonovae in the rise time-peak brightness phase-space is not as tightly constrained. They are still well separated by timescale but several of the contaminating sources have comparable peak brightnesses at 200 Mpc. As we consider more distant contaminant populations, we observe little to moderate reddening, with all of the contaminants remaining blue compared to the kilonovae population ($i-z\lesssim$ 0 mag to {\em z} $\sim$ 0.6). The $0.6+0.3\;M_{\odot}$ and $1.2+0.05\;M_{\odot}$ Type .Ia SNe models have $i-z\sim$ 0.2 mag, but their longer timescales $(t\sim\text{ week})$ keep them well separated from kilonovae. In addition, there is a systematic change in timescales, with all of the contaminants (except for the AIC events) shifting to a rise time of $\gtrsim  2$ d by {\em z} $\sim$ 0.6 due to time dilation.

\clearpage
The inherent low luminosity of most fast transients means that they quickly drop below the sensitivity limit of a reasonable search (e.g., $m_z \approx 23$ AB mag). Only the Type Ia SNe, Pan-STARRS fast transients, and the most luminous type .Ia models are brighter than $\approx$ 23 AB mag by {\em z} $\sim$ 0.2, and by {\em z} $\sim$ 0.6 only the transient PS1-11qr and Type Ia SNe will still be detectable.

@@

@@

>>>>

@@

\clearpage
Extensive theoretical and observational work has been performed to understand the potential EM emission from a merging compact object binary. The range of potential counterparts covers the entire EM spectrum. The most popular counterpart is a short-hard gamma-ray burst \citep[SGRB; e.g.,][]{Paczynski1986,Narayan+92,Berger2014} powered by rapid accretion onto the compact remnant which leads to a strongly beamed relativistic jet. The evidence for a connection between SGRBs and compact object mergers includes the association with older stellar populations than for long GRBs \citep{Berger+05,Bloom+06,Berger2011,Fong+11,Fong+13}, large physical offsets between the burst locations and their host galaxies indicative of progenitor kicks \citep{Fong+10, FongBerger13}, weak association with the underlying stellar light of their hosts \citep{Fong+10, FongBerger13}, and a lack of association with supernovae \citep{Fox+05,Soderberg+06,Berger2010,FongBerger13,Berger2014}. The relativistic jet can also be observed in the x-ray, optical, and radio as non-thermal afterglow emission arises from interaction with the ambient medium. This emission is also affected by relativistic beaming but may still be seen off-axis as the jet decelerates and expands. In particular, for large off-axis angles $(\theta\gtrsim 2 \theta_j)$ the emission will be most pronounced in the radio on timescales of weeks to years  \citep{NakarPiran11,MetzgerBerger12}.

@@

To investigate the detectability of kilonovae, and to asses potential contamination in deep high-cadence, wide-field searches for such sources, we carried out a set of observations using the Dark Energy Camera \citep[DECam][]{Flaugher+15} mounted on the CTIO/Blanco 4m telescope (PI: Berger). This GW follow-up ``dry-run" covered an area typical for a GW error region ($\sim$ 70 deg$^2$), with two visits per night for five nights to a depth of $\approx$ 24 AB mag in both {\em i}- and {\em z}-band. This rich data set will serve as the framework to study the nature of foreground/background contamination and to test the effectiveness of various observing strategies and contaminant rejection strategies when performing follow-up observations of a GW event.

@@

\section{Kilonovae and Contaminants}
\label{sec:ch2_contaminants}

@@

A critical property for understanding kilonova emission is the opacity, as the ability of photons to diffuse through the ejecta determines the timescale, peak luminosity, and the spectral energy distribution of the transient. \citet{Metzger+10}, using the models of \citep[LP98 hereafter]{LP98}, found that a typical Fe-peak element opacity $(\kappa \sim 0.1 \; \text{cm}^2 \text{ g}^{-1} \text{ for } \lambda \sim 0.3 - 1\; \mu\text{m})$ leads to a peak time of $t_p \sim 1$ day, with a bolometric peak luminosity of $L_p \sim 10^{41} \text{ erg s}^{-1}$, and a photospheric temperature of $T_p \approx 10^4$ K (i.e., a peak in the UV). However, subsequent work by \citet{Kasen+13} and \citet{BarnesKasen13} found that {\em r}-process material have much larger opacities with $\kappa \sim 10-100 \; \text{cm}^2 \text{ g}^{-1} \text{ for } \lambda \sim 0.3 - 3\; \mu\text{m}$. This leads to a transient that is longer lived $(t_p \sim$ days), fainter $(L_p \sim 10^{40}$ erg s$^{-1})$, and much redder $(T_p \approx 3000$ K, e.g., a peak in the NIR). These results have been confirmed in simulations by other groups \citep[see e.g.,][]{TanakaHotokezaka13,Grossman+14,Tanaka+14}. A discussion of the effects of modifications the effects of modifications to the standard kilonovae models on our results can be found in \cref{sec:ch2_altkilo}.

In this work, we construct kilonova light curves using the time-resolved spectra of \citet{BarnesKasen13}. They consider a grid of models with $\Mej = 10^{-3}\;, 10^{-2} \;, \text{ and } 10^{-1} \;M_{\odot}$ and with $\beta_{\text{ej}} = $ 0.1, 0.2, and 0.3. Unless otherwise noted we only consider the average ejecta properties of $\Mej =10^{-2}\;M_{\odot}$ and $\beta_{\text{ej}} = 0.2$ since we do not know the distribution of ejecta properties a priori. We compute light curves using the DECam filter transmission curves\footnote{\singlespace \url{http://www.ctio.noao.edu/noao/sites/default/files/DECam/DECam_filters_transmission.txt}}\footnote{\singlespace We note that the DECam transmission filters are fairly standard so these results are relevant for other instruments as well.} with the luminosity $L_{\phi}$, through a filter $\phi$, at a time t, being given by 
\begin{align}
\label{eq:LCint}
L_{\phi}(t) = \frac{\int d\lambda L_{\lambda}(t,\lambda) \phi(\lambda)}{\int d\lambda \phi(\lambda)}
\end{align}
\noindent where $L_{\lambda}(t,\lambda)$ is the model specific luminosity and $\phi(\lambda)$ is the filter transmission as a function of wavelength. A representative plot of the spectra at 1 day and 1 week, along with the resulting {\em r}-, {\em i}-, and {\em z}-band light curves are shown in \Cref{fig:ch2_kilosample}.

\begin{figure}[t!]
\centering
\includegraphics[width=\textwidth]{./figs/chapter2/f1.pdf}
\caption{\singlespace Top: Kilonova spectra from \citet{BarnesKasen13} at $\sim$ 1 d and $\sim$ 1 week with ejecta properties of $\Mej =10^{-2}\;M_{\odot}$ and $\beta_{\text{ej}} = 0.2$. Bottom: Light curve for the same model parameters computed at a fiducial distance of 200 Mpc in the DECam {\em r}, {\em i}- and {\em z}-band filters. \vspace{10pt}}
\label{fig:ch2_kilosample}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.7\textwidth]{./figs/chapter2/f2.pdf}
\caption{\singlespace Top: {\em i}-band light curves of the contaminants considered in this work at a fiducial distance of 200 Mpc. Bottom: {\em z}-band light curves. The SN2010U light curve has an offset applied for visibility. See \cref{sec:ch2_WDmerge,sec:ch2_fastnovae} for details.}
\label{fig:ch2_LCcont}
\end{figure}

@@

The slow evolution of the CSDD model light curve relative to kilonovae means that it can be easily removed from consideration as a contaminant on the basis of timescales alone. However, we still consider the ELDD scenario as the more rapid rise time and comparably faster decline means that it could serve as a potential contaminant. We make use of bolometric light curves from the simulations of \citet{Sim+12} to compute DECam {\em i}- and {\em z}-band light curves as in \cref{sec:ch2_WDmerge}. The ELDD spectrum is approximated as a blackbody with a characteristic temperature of $T_p \approx 4800$ K using the model spectra of \citet[see their Figure 7]{Sim+12}. The bolometric corrections are found to be $m_{\text{bol}} - m_{\text{i}} \approx -1.7$  mag and $m_{\text{bol}} - m_{\text{z}} \approx -1.9$ mag. We assume that these corrections are constant in time and apply them to the bolometric light curves to generate $i$- and $z$-band light curves (\Cref{fig:ch2_LCcont}).

@@

\subsection{Fast Classical Novae}
\label{sec:ch2_fastnovae}
Classical Novae (CNe) are close binary systems where a WD accretes hydrogen from a companion star that has filled its Roche lobe. If the accretion rate is sufficiently low then the hydrogen envelope on the WD will reach the critical temperature and density necessary to undergo thermonuclear runaway. The resulting energy deposition drives rapid expansion and mass loss in the surface layers of the unburnt hydrogen, powering the nova transient \citep{Warner2003}. The typical luminosity of such an event is $M_V \approx -7.5$ mag, but $\sim$1\% of nova reach a peak luminosity of $M_V \sim -10$ mag \citep{Shafter+09}. Despite their low luminosity compared to other transients considered in this work the high rate of CNe \citep[$\sim 35$ yr$^{-1}$ per MW/M31 type galaxy,][]{Darnley+06, Shafter+11} means they could be a significant foreground contaminant.

In this analysis, we consider the event SN2010U as our fiducial CN. It is an unusually luminous $(M_V \approx -10$ mag) and rapidly evolving $(t_{\text{decline}} \approx 3.5$ days) CN discovered in NGC 4214 \citep{NakanoKadota10,Czekala+13}. We utilize the $r$-band light curve presented in \citet{Czekala+13} as it is well-sampled and contains data points prior to maximum light. We model the spectrum of SN2010U as a blackbody with $T_p \approx 8000$ K as per the spectral fits in \citet{Czekala+13}. We integrate this spectrum over the DECam transmission filters to compute the $r-i$ and $r-z$ colors. We find $r-i \approx -0.43$ mag and $r-z \approx -0.85$ mag. We apply these colors to the $r$-band light curves from \citet{Czekala+13} to produce $i$- and $z$-band light curves (\Cref{fig:ch2_LCcont}).

\subsection{Stellar Flares}
\label{sec:ch2_flarestar}
Stellar flares, primarily from M dwarfs, are another potential contaminant in fast cadence wide-field optical searches \citep{Becker+04,KulkarniRau06,Berger+13}.  These flares have characteristic timescales of minutes to hours \citep[see e.g.,][]{Berger+13} and temperature of $T_p \sim 10^4$ K, leading to an appearance as fast and blue transients.  As a result, M dwarf flares will only appear as a significant contaminant in kilonova searches that rely on single-epoch detection, primarily in $g$- and $r$-band in which the kilonova timescale is shorter than in the redder bands.  We return to this point in \cref{sec:ch2_disc}.  The sky-projected M dwarf flare rate at the depths relevant for kilonova searches is large, $\sim 10$ deg$^{-2}$ day$^{-1}$ \citep{KulkarniRau06}.  On the other hand, as was demonstrated by \citet{Berger+13}, deep observations in the $i$- and $z$-band can easily reveal the underlying M dwarfs and hence reduce the potential for contamination from flares.

\subsection{AGN Variability}
\label{sec:ch2_agns}
We also consider short-duration optical variability from active galactic nuclei (AGNs). AGN typically show strong variability on month or year-long timescales, making AGN outbursts too long in duration to be confused with a kilonovae. However, recent work has detected rare cases of day-long  or shorter variability in the NIR. \citet{Genzel+03}  observed an H-band flare in Sgr A$^{*}$ during which the source brightened by a factor of 5. This flare lasted $\sim$ 30 min, with a characteristic rise time of $\sim2-5$ min. Such short timescale flares can be ruled out as kilonova contaminants by requiring a detection on more than a single night. Flares with longer durations are also unlikely to be significant contaminants. \citet{Totani+05} performed {\em V}- and {\em i}-band observations of six AGN, selected on the basis of their variability, over a period of four days. They found typical variability of only $\sim1-5\%$ above their host galaxy luminosity. More recent work by \citet{Ruan+12} studied the variability of a sample of \fermi\ selected blazars. They found that these sources showed a typical rms variability of $\sim0.6$ mag on timescales of $\sim 10$ d. Therefore, AGN are unlikely to exhibit sufficiently large variability to be confused with a kilonova.

\subsection{Volumetric Rates}
\label{sec:ch2_rates}
An additional important factor for understanding the impact of potential contaminants on fast cadence, wide-field searches is their volumetric rates. We note that the rates for these types of objects are generally poorly understood due to the lack of strong observational constraints; the notable exceptions are Type Ia SNe and fast novae. The volumetric rates of potential contaminants are compiled in \cref{tab:ch2_rates}. No values are provided in the literature for the ELDD volumetric rate so we assign a fiducial value of 1\% of the Type Ia SN rate, consistent with the other rapid transients in this study. The volumetric rate for Type Ia SNe is known to be a function of redshift, and we use an approximate average volumetric rate in redshift bins of $\Delta z \sim 0.25$ out to $z_{\text{max}} \sim 0.75$. \citep[][and references therein]{Li+11,Pereira+13,Graur+14}. These rates are listed in \cref{tab:ch2_rates_Iae}.

\singlespace
\begin{deluxetable}{cccccc}
\tabletypesize{\footnotesize}
\tablecolumns{6} 
\tabcolsep0.05in\footnotesize
\tablewidth{0pt} 
\tablecaption{Expected Rates For Fast Transients}
\tablehead{
\colhead{Object}&  \colhead{$\mathcal{R}_{\rm vol}$}& \colhead{$z_{\rm max}$}  &  \colhead{$\mathcal{R}_{\rm area}$} &  \colhead{$N$}  &  \colhead{Reference}\\
\colhead{} & \colhead{(Mpc$^{-3}$ yr$^{-1}$)} & \colhead{} & \colhead{(deg$^{-2}$ yr$^{-1}$)} &  \colhead{} & \colhead{} }
\startdata
 Type .Ia &10$^{-6}$ &  0.5  &  $0.7$&1.4  & \citet{Bildsten+07} \\
 WD-NS mergers  &10$^{-5}$ &0.06  & $2\times10^{-2}$  &$3\times10^{-2}$  &  \citet{Thompson+09}  \\
 WD-BH mergers  &10$^{-5}$ & 0.06  &$2\times10^{-2}$  &$3\times10^{-2}$  & \citet{fryer+99} \\
AIC &10$^{-6}$ & 0.06  & $2\times10^{-3}$&  $3\times10^{-3}$ &  \citet{Darbha+10} \\
ELDD&3$\times 10^{-7}$ & 0.14  & $6\times10^{-3}$  &$1\times10^{-2}$ &  $\cdots$  \\
 Pan-STARRS fast &5$\times 10^{-6}$ &  0.5  &$3.5$&  7 & \citet{Drout+14}  \\
 Fast CNe & 3.5$\times 10^{-4}$ & 0.009 & 2$\times 10^{-3}$ & 4$\times 10^{-3}$ & \citet{Darnley+06}
 \enddata
\tablecomments{Expected rates for the various contaminants considered in \cref{sec:ch2_contaminants}. $\mathcal{R}_{\rm area}$ is computed assuming an isotropic distribution of sources in a volume defined by the comoving volume at $z_{\rm max}$.  The column $N$ refers to the number of events expected during a search covering 100 deg$^2$ for 7 days. See \cref{sec:ch2_rates} or details.}
\label{tab:ch2_rates}
\end{deluxetable}

\begin{deluxetable}{cccc}
\tabletypesize{\footnotesize}
\tablecolumns{4} 
\tabcolsep0.05in\footnotesize
\tablewidth{0pt} 
\tablecaption{Expected Rates For Type Ia SNe}
\tablehead{
 \colhead{Redshift}  &  \colhead{$\mathcal{R}_{\rm vol}$}&  \colhead{$\mathcal{R}_{\rm area}$}  &  \colhead{$N$} \\
 \colhead{} & \colhead{(Mpc$^{-3}$ yr$^{-1}$)} &  \colhead{(deg$^{-2}$ yr$^{-1}$)} &  \colhead{} }
\startdata
0 -- 0.25 & $3\times10^{-5}$ & 3.5 & 21  \\
0.25 -- 0.50 & $4\times10^{-5}$ & 25 & 153  \\
0.5 -- 0.75 & $5\times10^{-5}$ & 65 & 394  
\enddata
\tablecomments{Expected rates for Type Ia SNe in redshift bins of $\Delta z = 0.25$ out to $z_{\rm max} = 0.75$. See \cref{sec:ch2_rates}, \citet{Graur+14}, and references therein.}
\label{tab:ch2_rates_Iae}
\end{deluxetable}
\doublespace

To estimate the expected number of events observed during a GW optical follow-up search we compute the number of sources per square degree per year. This is done by determining the maximum redshift to which the object can be observed to a limiting magnitude of $z \approx$ 23 mag (see \cref{sec:ch2_kilo,sec:ch2_MCsims_det}). This rate is then multiplied by the search area $(A \sim 100$ deg$^2$) and duration to get the number of expected sources for our observations. We compute the duration by assuming that a kilonova will be detected within approximately one day of the GW trigger. Therefore, we do not consider events that begin more than +2 d into the search. However, we do consider sources that begin up to twice their characteristic timescale ($\sim5$ d for rapid transients and $\sim 20$ d for Type Ia SNe) prior to the GW trigger, as they may remain visible when the search begins. The total duration is then 7 d for rapid transients and 22 d for Type Ia SNe. The number of expected sources is given in \cref{tab:ch2_rates,tab:ch2_rates_Iae}. Type Ia SNe dominate the number of expected events with $N \sim 568$ when summing across all redshift bins. Of the population of fast transients, only type .Ia and the Pan-STARRS fast transients are expected to have $N \gtrsim 1$. Despite the low expected occurrence rate for other fast transients, we still consider these sources in this analysis due to the large uncertainty in their rates.

@@

\begin{figure}[t!]
\centering
\includegraphics[width=\textwidth]{./figs/chapter2/f3.pdf}
\caption{\singlespace Time-color (left) and time-magnitude (right) slices of the time-color-magnitude phase-space at a selection of redshifts. Arrow markers denote upper limits on the rise time. Note that in all panels the population of kilonovae are kept at the fiducial distance of $\sim$ 200 Mpc while redshift corrections are applied to the contaminant population. Kilonovae are well separated from the contaminant population in both color and timescales. The delineation in magnitude is less apparent. This figure does not include CNe as they are too faint to be relevant background contamination.}
\label{fig:ch2_phase}
\end{figure}

\clearpage
Investigating the first panel in the rise time-color space, we can see that the nine kilonovae models are well separated from the population of contaminating sources by virtue of their redder color. We also note that their rise times are generally shorter, with only the AIC and $1.0+0.05\;M_{\odot}$ type .Ia light curves exhibiting comparable timescales. The region occupied by kilonovae in the rise time-peak brightness phase-space is not as tightly constrained. They are still well separated by timescale but several of the contaminating sources have comparable peak brightnesses at 200 Mpc. As we consider more distant contaminant populations, we observe little to moderate reddening, with all of the contaminants remaining blue compared to the kilonovae population ($i-z\lesssim$ 0 mag to {\em z} $\sim$ 0.6). The $0.6+0.3\;M_{\odot}$ and $1.2+0.05\;M_{\odot}$ Type .Ia SNe models have $i-z\sim$ 0.2 mag, but their longer timescales $(t\sim\text{ week})$ keep them well separated from kilonovae. In addition, there is a systematic change in timescales, with all of the contaminants (except for the AIC events) shifting to a rise time of $\gtrsim  2$ d by {\em z} $\sim$ 0.6 due to time dilation.

The inherent low luminosity of most fast transients means that they quickly drop below the sensitivity limit of a reasonable search (e.g., $m_z \approx 23$ AB mag). Only the Type Ia SNe, Pan-STARRS fast transients, and the most luminous type .Ia models are brighter than $\approx$ 23 AB mag by {\em z} $\sim$ 0.2, and by {\em z} $\sim$ 0.6 only the transient PS1-11qr and Type Ia SNe will still be detectable.

@@

\begin{figure}[t!]
\centering
\includegraphics[width=\textwidth]{./figs/chapter2/f11.pdf}
\caption{\singlespace The decline time-$\Delta m_z$ slice of the decline time-color-magnitude phase-space. The slow decline of Type Ia SNe and the Pan-STARRS fast transients separate them from the rapidly evolving kilonovae. However, the type .Ia and the miscellaneous fast transients show changes in brightness comparable to those of kilonovae.}
\label{fig:ch2_decdm}
\end{figure}

@@

\item In the case where a pre-existing template is not available, cuts on $i-z$ color and timescale are still effective for separating kilonovae from the contaminant population for sources observed during the rise to peak (\cref{sec:ch2_diff}). However, for sources seen only in decline, these cuts are significantly less effective. We do note that by utilizing a detection criterion based on the change in brightness in an observed source we can reduce the number of contaminant sources detected. However, this also reduces the kilonovae detection rates. 
